\section{Conclusion}

\subsection{Synthesis}
\begin{frame}{Key Contributions}
    \begin{itemize}
        \item Identified vulnerabilities in SoC buses (Wishbone, AXI-Lite, AXI).
        \item Developed experiment framework using LiteX + VerifyPin + FISSA.
        \item Demonstrate limitation of software countermeasure
        \item Proposed hardware-based countermeasure for the Wishbone bus.
        \item Demonstrated full protection under evaluated fault models.
    \end{itemize}
\end{frame}

\begin{frame}{Experimental Highlights}
    \begin{itemize}
        \item Over 40 million simulations executed on 3 Xeon Gold servers. 
        \item Total campaign duration: more than 2 months.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Global Reflection}
\begin{frame}{Global Reflection}
    \begin{itemize}
        \item Bus-level vulnerabilities exist across Wishbone, AXI-Lite, AXI.
        \item Real-world SoCs lack built-in protection for handshake signals.
        \item Multi-bit fault injection remains a realistic threat with advanced equipment.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Perspectives}
\begin{frame}{Future Directions}
    \begin{itemize}
        \item Apply countermeasures to AXI-Lite and AXI buses.
        \item Combine advanced protections (CFI, runtime checks) against both software/hardware attack.
        \item Integrate protection into LiteX to add the security dimension.
        \item Evaluate NoC-based architectures under fault injection attacks.
    \end{itemize}
\end{frame}